\section{Introduction}
\addcontentsline{toc}{chapter}{Introduction}


% Set page numbering to arabic the first time we commence a chapter.
% This is required to get the page numbering correct.
\pagenumbering{arabic}
\doublespace{}

I propose to explore a range of extended techniques that utilise the harmonic series and to assess how they can be used in my, and other people's, creative practice in composition and performance. 
This exegesis considers the techniques associated with \underline{\textcolor{blue}{\href{https://vimeo.com/57219720}{subharmonics}}}, \underline{\textcolor{blue}{\href{https://vimeo.com/81380931}{multiphonics}}}, and \underline{\textcolor{blue}{\href{https://vimeo.com/92221539}{half-harmonics}}} on string-based instruments.\footnote{These video examples are part of Ellen Fallowfield's CelloMap project. I chose to use them as examples because they are performed by the same professional musician, on the same instrument, in a professional studio, which helps to establish a baseline of what can be expected of the techniques.}
For purposes of brevity, these harmonic-based extended techniques will simply be referred to as `techniques' throughout the paper, except for when differentiation between standard techniques is needed.

While some techniques such as harmonics are well established and understood, others, such as subharmonics, are still underdeveloped in terms of both repertoire and resources available. 
The timbral potentials of these techniques are uncharted territories and collectively represent an entire sound world that remains relatively inaccessible to composers.
% TODO: Evidence for uncharted territories - https://trello.com/c/u605dDsM/15-evidence-for-uncharted-territories

% To identify where further research is required, I will conduct a review of the literature and resources that are readily available to composers to assess what techniques require further investigation and refinement. 
% By researching these techniques and the mechanics behind them, interviewing professionals, and analysing recordings made, I hope to gain a better understanding of how these techniques can be implemented in my practice. 
% As part of both the analysis of techniques and my compositional practice, I will assess not only the compositional potential, but also the practicality of techniques. 
% Reviewing the feasibility and notational aspects of the techniques will render the exegesis a practical document to reference for performance and composition.


The exegesis is intended to serve as a useful reference source for artists interested in learning about the mechanics, qualities, and potential implementations of these harmonic based extended techniques. 
The works that I compose accompanying the exegesis will show idiomatic treatment of the techniques and serve as references as such in the exegesis.
This exegesis makes use of hyperlinks throughout to enhance its usefulness as a reference document, and ease of sharing.
The dissemination of the material I research will contribute to the accessibility of new sound possibilities for artists.

% The folio of completed works is: \begin{itemize}
    % \item \underline{\textcolor{blue}{\href{https://soundcloud.com/rhysgray/what-are-you-doing-with-the-humans/s-PLerb}{\violinPiece}}}
    % \item \underline{\textcolor{blue}{\href{https://soundcloud.com/rhysgray/doppelganger/s-QV34l}{\violaPiece}}}
    % \item \underline{\textcolor{blue}{\href{https://rhysgraymusic.com/celloPieceHonours}{\celloPiece}}}
% \end{itemize} 
% what they are
% what the field
% Why i'm interested
% Why I want to do it
% Why its relevant
% To integrate it into the field and wheelhouse

