
\subsection{Methodology}
% Autoethnography, document analysis, interviews, research
Through interviews with players at varying stages of proficiency and familiarity with the techniques, I will be able to uncover the barriers to producing these techniques. 
Document analysis of existing resources and compositions will help direct and support the line of enquiry. 
Autoethnography of my creative process will document the research process and clarify my intent.  

To understand extended techniques, we must first establish a basis of what is considered a `regular' technique, or rather, what the qualifying factors are for a technique to be considered extended.
To do this, we will look at what techniques are commonplace in the literature, and which are less so.
Techniques that require descriptions in the frontmatter would reasonably be understood to be considered extended techniques.

The aim of this research project is not to make the techniques popular enough to make clarification of technique unnecessary, or for it to enter the canon of techniques so that it is no longer considered to be `extended' (as the Bartok pizzicato has).
Rather, this is intended to act as a resource for composers and artists to be drawn upon as a reference for when they wish to use the technique.
A considered and informed judgement call over a technique can only be made when the technique is understood well.
The composer will communicate the information necessary to realise the technique to the player, typically through the frontmatter. 
In order to better understand what information composers deem useful to communicate to players, a review of scores with similar techniques will take place.
By breaking the score's frontmatter content up into its actions, we can understand how composers communicate their desired techniques to players.
